% !TeX root = ../main.tex
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Introduction
\chapter{Introduction}
\label{chap:introduction}

This thesis aims at exploring a possible solution for automation in the testing of robotic systems through the medium of a domain-specific language (DSL) and simulation software.

This chapter intends to introduce the motivation for this work (Section 1.1), present the problem statements of such an approach (Section 1.2), discuss the objectives (Section 1.3), present the expected contributions (Section 1.4), and finally summarize the structure of the rest of the document (Section 1.5).


% ------------------------------------------------------------------------------
% Motivation
\section{Motivation}
\label{sec:motivation}

Robotics already significantly impact our current society, industrially, for instance, in medicine and agriculture, or leisurely in contests or personal use. Robotics often take critical roles like the example of robot arms in car assembly lines or autonomous farms. The tendency is for robot usage to keep growing at a global level. 

Robotic Systems are non-deterministic, mainly because robots interact directly with the real world. Testing software in such environments is complex, as many variables can change, and verifying the success of a task or movement may not be possible from the robot's perspective, and external monitoring may be required.

Current practices in testing robot software mainly involve field testing, simulation testing, and log checking and require a human to analyze the robot's behavior to determine whether the behavior is correct. Due to their broad practicality, the quality of software running on robots should be extremely important to us. Robot software, as well as the techniques used to test their quality, are very field-specific and different from the techniques employed in traditional Software Engineering mainly because of their real-world interaction. This peculiarity means automatic tests are barely used in robotics. Studying possible options for viable automation of tests in robotic systems could lead to an opening on its usage in both research and the industry. Also, allowing for multiple parallel executions of tests not depending on human monitoring could improve the quality of current and future robot software.


\section{Problem Statement}
\label{sec:problem}

The multiple challenges in robot testing influence planning how to test a robot because there are tradeoffs among choices.

Using simulation-based testing, the developers can take advantage of real values of objects' attributes to compare with what the robot system perceives. Using this alleyway, it is possible to, in a way, surpassing the need for human-in-the-loop testing. 
 
While simulation-based tests are a promising approach for automation, there is still distrust in the precision and validity of the results. As a result, simulation-based autonomous testing is barely used due to reliability and factors like cost and complexity. Due to these factors, despite being dangerous, sometimes expensive, or work-intensive, real-life robot testing or other methods are still the main choices. The resulting product is a lack of quality in the software across projects.

When developing a DSL for simulation-based autonomous tests, problems like what components to monitor and how to express them arise. Having a DSL to specify a robotic system's properties can be useful. However, there is a need to control its complexity and accessibility, or it can become a burden in the testing process.


% ------------------------------------------------------------------------------
% Objectives
\section{Objectives}
\label{sec:objectives}

The ultimate goal of this thesis is to remove the need for human-in-the-loop testing of robotic systems by studying a possible solution for automation in simulation-based tests.

This work aims to provide developers with a way to verify their robotic systems' temporal and positional properties automatically. We propose the introduction of a DSL for developers to express their relevant properties. The given properties compile into monitors that can be used in simulation to ensure the correctness of the system. The DSL was designed from the point of view of the Robot Operating System (ROS)~\cite{quigley2009ros} developers and tries to abstract the underlying Linear Temporal Logic (LTL) system. The DSL allows properties to reason about native ROS constructs, like \textit{topics}, \textit{messages} and simulation information. Thus, it is possible to express properties that relate the internal information of the system with the corresponding information in the simulator.

The DSL should allow describing a robotic system's properties simply and intuitively while simultaneously expressing relevant temporal and positional arguments between robots components and objects in the simulation.


\section{Contributions}
\label{sec:contributions}

The expected contributions of this thesis are below enumerated.

\begin{enumerate}
    \item Definition of a domain-specific language to specify robotic systems' properties.
    \item Implementation of a compiler for the language that can generate software capable of monitoring relevant components while in a simulation.
    \item Evaluation of the expressive capabilities of the solution.
\end{enumerate}

% ------------------------------------------------------------------------------
% Structure of the document
\section{Structure of the document}
\label{sec:structure}

The document is organized as follows:

\begin{itemize}
    \item Chapter 2 - Background \& Related Work:
    \item Chapter 3 - Motivational Example
    \item Chapter 4 - Specification Language for Robotics Properties
    \item Chapter 5 - DSL Usage Examples
    \item Chapter 6 - Monitoring
    \item Chapter 7 - Evaluation 
    \item Chapter 8 - Future Work
    \item Chapter 9 - Conclusion
\end{itemize}
